Modular transmission line probes for microfluidic nuclear magnetic resonance spectroscopy and imaging

### 2019-02-05 ms5n17-tlp-mn-180216-001.tex
Put in the final figure

In the remainder of this paper, we discuss the probe design in more detail, 
and characterise its performance in terms of sensitivity and resolution. 
We also demonstrate its application in microfluidic cell culture, 
protein NMR spectroscopy, and micro-imaging. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Probe Design}
\label{sec:probe-design}
%Figure showing modular probe concept; example of probe base with several inserts\newline
%\begin{itemize}
%	\item Philosophy: use cheap materials and prefab Al profiles wherever possible
%	\item Hold it together with a small number of precision-machined parts
%	\item Generic probe base design that can accommodate a variety of inserts
%	\item Insert: PCB that accommodates both tuning/matching circuitry for up to two channels and the transmission line probe itself
%	\item Tuning and matching capacitors for each channel are part of the probe base
%	\item Connection using standard coaxial RF connectors (SMP)
%	\item Microfluidic chip design; outline
%	\item Design/explosion drawings
%	\item Circuit design for double resonance
%	\item Transmission line connection between tuning/matching of 1H channel and resonator (Smith chart argument)
%	\item 38 mm bore (Bruker NB) and 45 mm bore (Varian/Agilent NB)
%\end{itemize}
\begin{figure}
\centering
\includegraphics[width=.5\linewidth,keepaspectratio=true]{./figures/ms5n17-tlp-im-181218-probe.png} 
\caption{The probe with and without sleeve.}
\label{fig:probe} 
\end{figure}
The probe assembly design is based on readily and easily available materials and parts 
with minimal need for customised fabrication.
All mechanical and electronic components are readily available from commercial sources, 
with the exception of a small number of precision-machined parts which are used 
to hold the assembly together. These can be ordered from commercial prototyping 
manufacturers or machined in-house if mechanical workshop facilities are available.

\begin{figure}
\centering
\includegraphics[width=.7\linewidth,keepaspectratio=true]{./figures/ms5n17-tlp-im-181007-Probe-explode.png} 
\caption{Exploded view of the probe design (left) and assembled probe (right). 
Probe base (blue) and scaffold assembly (green) with TMPCB (brown, 8) provides a 
generic mounting point for the NMR detector assembly (red). 
A microfluidic device (1) can be inserted from the top of the probe. 
The device is held in place by a spacer (4) between the detector planes (3 and 5) 
ensuring alignment of the sample chamber with the sensitive area on the PCB planes. 
The X channel frequency is controlled by a 
modular PCB insert (6). The modular NMR detector is made up of parts 3 to 6 and is 
connected to the TMPCB (8) through standard coaxial (SMP) connectors. 
The scaffold assembly is made up of aluminium tubes (11), scaffold support disks (10) 
and, auxiliary support disks (12).}
\label{fig:probe-explode} 
\end{figure}

@@

The scaffold assembly is made up of two hollow aluminium tubes (11) held together with 
precision machined parts at both ends (scaffold support disks (10)). 
The semi-rigid cables are  threaded through these hollow aluminium tubes 
to reach the detector assembly. 
Auxiliary support disks (12) provide additional rigidity for the scaffold assembly. 
In our case, these were laser cut from 5~mm thick PMMA sheets, 
but they may also be made by conventional machining.
A tuning and matching PCB (TMPCB, 8) is held in place on the top of the probe skeleton by
the detector support rods (9), with the semi-rigid coaxial cables providing
additional support.
The TMPCB is permantly fixed to the probe base, and serves as attachment point
for the exchangeable detector assembly. It also provides 
capacitors (7) for secondary tuning and matching of the detector.
The detector assembly is connected to the TMPCB through a pair of
low-loss SMP connectors (14 in \fig{fig:ProbePhoto}).
While each detector assembly is roughly
tuned and matched with fixed circuit elements soldered onto it,
the adjustable capacitors on the TMPCB are used for 
fine tuning and matching 
at the magnet through tuning rods (13) which reach all the way to the probe base.

The base, the scaffold assembly and the tuning and matching PCB all have a free circular 
axial bore providing access from the bottom of the probe, which can be used for 
temperature and environmental control and/or fluidic connections to the sample as
needed.
The detector assembly is made up of two PCBs (3 and 5), a spacer between them (4) and an 
RF-insert (6). The detector used here is based on a planar transmission line geometry 
as described elsewhere~\cite{gream_2016,stripline_jan}; other structures such as spiral
coils and Helmholtz pairs are also possible, and can easily be realised on a PCB.
The transmission line detector consists of a pair of copper 
planes, each with a constriction at the centre.  This geometry gives rise to an 
electromagnetic eigenmode with a strong anti-node of the magnetic field between
the two planes at the site of the constriction. This concentrates the rf field 
and the detection sensitivity
onto the sample area. 
The shape and size of the copper structures have been chosen such as to produce
an eigenfrequency of this mode around 630~MHz. The PCBs have been designed with
solder points at the back side for chip capacitors; these can be used to tune the
resonance down to the desired proton Larmor frequency (500 or 600~MHz in the present case).
Both the PCBs (3 and 5) also carry circuitry for channel separation and primary 
matching of the detector. A copper or brass auxiliary disk (2) holds the PCBs 
together at the top, and forms the top end of the probe assembly.
The probe sheath (not shown in~\fig{fig:probe-explode}) is fixed to the auxiliary disk (2) 
and the scaffold assembly with screws. 
An exchangeable microfluidic device (1) can be inserted from the top of the detector. 
Microfluidic devices can be used as passive sample holders, filled before putting them
into the probe,
or the device can be actively perfused and operated in-situ. 
The microfluidic device is held in place by the spacer 
between the two PCB planes coinciding the sample chamber with the constriction on the 
detector.  The 45-degree edges of the device outline help with precise and reliable
alignment. 
Since the detector assembly is attached to the probe base through SMP 
connectors, it can easily be exchanged. 
Using standard PCB processes to manufacture the detector provides great flexibility
to optimise detector performance for different applications, including different
microfluidic sample sizes, and different combinations of target nuclei.

\begin{figure}
\centering
\includegraphics[width=.5\linewidth,keepaspectratio=true]{./figures/ms5n17-tlp-im-181007-both-detectors.png}
\caption{Dual channel probes with (A) and without (B) modular RF insert (6) for X channel. 
O.D. of A is 44.45 mm and B is 38 mm. Circuit elements of X channel are integrated on the 
PCB in B making it compact. Detachable SMP connectors (14) are used to 
connect the RF insert (6) and the detector assembly.}
\label{fig:ProbePhoto}
\end{figure}
\fig{fig:ProbePhoto} shows two different versions of the probe with and without the RF insert (6) 
for the X channel. In (A) the X channel can be tuned to different nuclei by exchanging 6. 
However, in (B) the circuitry of the RF insert (6) is integrated on the PCB plane (5), 
making the design more compact.
 Probe (A) is compatible with a Varian 600 Premium shielded 14.1 T magnet, with an outer diameter (O.D.) of 44.45 mm including the sheath. Probe (B), with an 
O.D. of 38 mm including the sheath, is designed to fit inside a 
Bruker narrow bore shim stack. 
Probe (B) also fits inside a Bruker micro-imaging gradient unit. Naturally, B can also be used with 
wide bore magnets with spacers or gradient stack to keep the probe centred.

>>>>

In the remainder of this paper, we discuss the probe design in more detail, 
and characterise its performance in terms of sensitivity and resolution. 
We also demonstrate its application in microfluidic cell culture, 
protein NMR spectroscopy, and micro-imaging. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Probe Design}
\label{sec:probe-design}
%Figure showing modular probe concept; example of probe base with several inserts\newline
%\begin{itemize}
%	\item Philosophy: use cheap materials and prefab Al profiles wherever possible
%	\item Hold it together with a small number of precision-machined parts
%	\item Generic probe base design that can accommodate a variety of inserts
%	\item Insert: PCB that accommodates both tuning/matching circuitry for up to two channels and the transmission line probe itself
%	\item Tuning and matching capacitors for each channel are part of the probe base
%	\item Connection using standard coaxial RF connectors (SMP)
%	\item Microfluidic chip design; outline
%	\item Design/explosion drawings
%	\item Circuit design for double resonance
%	\item Transmission line connection between tuning/matching of 1H channel and resonator (Smith chart argument)
%	\item 38 mm bore (Bruker NB) and 45 mm bore (Varian/Agilent NB)
%\end{itemize}
\begin{figure}
\centering
\includegraphics[width=.5\linewidth,keepaspectratio=true]{./figures/ms5n17-tlp-im-190205-rf-device-assembly.png} 
\caption{(A) Orientation of the device with respect to the constriction on the detector (PCB plane 5 is hidden). (B) Exploded view of the rf assembly with the device. }
\label{fig:rf-device-assembly} 
\end{figure}

\begin{figure}
\centering
\includegraphics[width=.5\linewidth,keepaspectratio=true]{./figures/ms5n17-tlp-im-181218-probe.png} 
\caption{The probe with and without sleeve.}
\label{fig:probe} 
\end{figure}
The probe assembly design is based on readily and easily available materials and parts 
with minimal need for customised fabrication.
All mechanical and electronic components are readily available from commercial sources, 
with the exception of a small number of precision-machined parts which are used 
to hold the assembly together. These can be ordered from commercial prototyping 
manufacturers or machined in-house if mechanical workshop facilities are available.

\begin{figure}
\centering
\includegraphics[width=.6\linewidth,keepaspectratio=true]{./figures/ms5n17-tlp-im-181007-Probe-explode.png} 
\caption{Exploded view of the probe design (left) and assembled probe (right). 
Probe base (blue) and scaffold assembly (green) with TMPCB (brown, 7) provides a 
generic mounting point for the NMR detector assembly (red). 
A microfluidic device (1) can be inserted from the top of the probe. 
The device is held in place by a spacer (4) between the detector planes (3 and 5) 
ensuring alignment of the sample chamber with the sensitive area on the PCB planes.
The modular NMR detector is made up of parts 3,4 and, 5 and is 
connected to the TMPCB (7) through standard coaxial (SMP) connectors. 
The scaffold assembly is made up of aluminium tubes (11), scaffold support disks (9) 
and, auxiliary support disks (12).}
\label{fig:probe-explode} 
\end{figure}

@@

The scaffold assembly is made up of two hollow aluminium tubes (11) held together with 
precision machined parts at both ends (scaffold support disks (9)). 
The semi-rigid cables are  threaded through these hollow aluminium tubes 
to reach the detector assembly. 
Auxiliary support disks (12) provide additional rigidity for the scaffold assembly. 
In our case, these were laser cut from 5~mm thick PMMA sheets, 
but they may also be made by conventional machining.
A tuning and matching PCB (TMPCB, 7) is held in place on the top of the probe skeleton by
the detector support rods (8), with the semi-rigid coaxial cables providing
additional support.
The TMPCB is permantly fixed to the probe base, and serves as attachment point
for the exchangeable detector assembly. It also provides 
capacitors (6) for secondary tuning and matching of the detector.
The detector assembly is connected to the TMPCB through a pair of
low-loss SMP connectors (14 in \fig{fig:ProbePhoto}).
While each detector assembly is roughly
tuned and matched with fixed circuit elements soldered onto it,
the adjustable capacitors on the TMPCB are used for 
fine tuning and matching 
at the magnet through tuning rods (10) which reach all the way to the probe base.

The base, the scaffold assembly and the tuning and matching PCB all have a free circular 
axial bore providing access from the bottom of the probe, which can be used for 
temperature and environmental control and/or fluidic connections to the sample as
needed.
The detector assembly is made up of two PCBs (3 and 5), a spacer between them (4). The detector used here is based on a planar transmission line geometry 
as described elsewhere~\cite{gream_2016,stripline_jan}; other structures such as spiral
coils and Helmholtz pairs are also possible, and can easily be realised on a PCB.
The transmission line detector consists of a pair of copper 
planes, each with a constriction at the centre.  This geometry gives rise to an 
electromagnetic eigenmode with a strong anti-node of the magnetic field between
the two planes at the site of the constriction. This concentrates the rf field 
and the detection sensitivity
onto the sample area. 
The shape and size of the copper structures have been chosen such as to produce
an eigenfrequency of this mode around 630~MHz. The PCBs have been designed with
solder points at the back side for chip capacitors; these can be used to tune the
resonance down to the desired proton Larmor frequency (500 or 600~MHz in the present case).
Both the PCBs (3 and 5) also carry circuitry for channel separation and primary 
matching of the detector. A copper or brass auxiliary disk (2) holds the PCBs 
together at the top, and forms the top end of the probe assembly.
The probe sheath (not shown in~\fig{fig:probe-explode}) is fixed to the auxiliary disk (2) 
and the scaffold assembly with screws. 
An exchangeable microfluidic device (1) can be inserted from the top of the detector. 
Microfluidic devices can be used as passive sample holders, filled before putting them
into the probe,
or the device can be actively perfused and operated in-situ. 
The microfluidic device is held in place by the spacer 
between the two PCB planes coinciding the sample chamber with the constriction on the 
detector.  The 45-degree edges of the device outline help with precise and reliable
alignment. 
Since the detector assembly is attached to the probe base through SMP 
connectors, it can easily be exchanged. 
Using standard PCB processes to manufacture the detector provides great flexibility
to optimise detector performance for different applications, including different
microfluidic sample sizes, and different combinations of target nuclei.

\begin{figure}
\centering
\includegraphics[width=.5\linewidth,keepaspectratio=true]{./figures/ms5n17-tlp-im-181007-both-detectors.png}
\caption{Dual channel probes with (A) and without (B) modular RF insert (13) for X channel. 
O.D. of A is 44.45 mm and B is 38 mm. Circuit elements of X channel are integrated on the 
PCB in B making it compact. Detachable SMP connectors (14) are used to 
connect the RF insert (13) and the detector assembly.}
\label{fig:ProbePhoto}
\end{figure}
\fig{fig:ProbePhoto} shows two different versions of the probe with and without the RF insert (13) 
for the X channel. In (A) the X channel can be tuned to different nuclei by exchanging 13. 
However, in (B) the circuitry of the RF insert (13) is integrated on the PCB plane (5), 
making the design more compact.
 Probe (A) is compatible with a Varian 600 Premium shielded 14.1 T magnet, with an outer diameter (O.D.) of 44.45 mm including the sheath. Probe (B), with an 
O.D. of 38 mm including the sheath, is designed to fit inside a 
Bruker narrow bore shim stack. 
Probe (B) also fits inside a Bruker micro-imaging gradient unit. Naturally, B can also be used with 
wide bore magnets with spacers or gradient stack to keep the probe centred.
